Cover letter for the IJRR submission \\ ``An efficient acyclic contact planner for multiped robots''

### 2016-01-21 body/overview.tex
Work together with Steve to review up to Section 2

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed in a sequential fashion: we first plan a guide path with the \textit{reachability condition}, before
extending it into a sequence of contact configurations in static equilibrium.
In this Section we describe this workflow and give some notations used throughout the paper. Some key concepts are introduced,
the definitions of which can be found in the glossary at the end of this paper.

@@

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a relevant guide path. The objective is to compute a path of root configurations that will allow contact creation. A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of contact configurations in static equilibrium.
The sets we introduce are illustrated in Figure~\ref{fig:dedefeas}.

\begin{figure}[t]
\centering
  \begin{overpic}[width=1\linewidth]{figures/2D_feas}
		\put (15,) {$C_{Equil}^0$      $\subset$} 
		\put (47,) {$C_{Contact}^0$ $\approx$ } 
		\put (76,) {$C_{Reach}^0$} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Illustration of several root configurations sets used in this paper in a 2D scene. Obstacles are violet, and units are in meters. To show the sets in a 2D representation, all the rotational joints of HRP-2 are locked in the shown configuration, such that a torso configuration
is only described by two positional parameters (x and y). $C_{Equil}^0$ is included in $C_{Contact}^0$. $C_{Reach}^0$ approximates $C_{Contact}^0$. Depending on a parametrization, we can obtain $C_{Contact}^0 \subset C_{Reach}^0$. Considering the configurations around the top obstacle, we can observe a dramatic
divergence between  $C_{Equil}^0$  and $C_{Contact}^0$ when the problem is not \textit{\gls{cluttered}}.}
		   \label{fig:dedefeas}
\end{figure}

More precisely, a root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that the robot satisfies two conditions:
\begin{description}
\item[C1]: is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: is in static equilibrium. 
\end{description}
A requirement for C2 is the condition C3: at least one end-effector must be in contact with the environment.
C1 and C3 define \glslink{contact feasible}{\textit{contact feasibility}}, and can be efficiently approximated even without computing the contact points. C2 defines \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}, that we describe with the set $C_{Equil}^0$ (Figure~\ref{fig:dedefeas} -- Green). C2 is difficult to approximate and requires the computation of the contact points, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.

We make the assumption that for the \textit{\gls{cluttered} contact planning problem} (that we define as admitting a solution where every configuration contains at least one contact
is \gls{quasi-flat}) verifying C2 is not critical for $\mathcal{P}_1$.
This means that root configurations satisfying C1 and C3 also satisfy C2. For this reason we neglect C2 during the resolution of \Pa, and we consider it only when solving \Pb.

We define $C_{Contact}^0$ (Figure~\ref{fig:dedefeas} -- Orange) as the set of \gls{contact feasible} root configurations (that can satisfy C1 and C2).
Planning in $C_{Contact}^0$ boils down to planning in $\mathbb{R}^r$, which has an acceptable practical complexity.

An intuitive description of \gls{contact feasible} configurations is ``close, but not too close'': close, because a contact surface must be partially included in the reachable workspace of the robot (represented for the right leg in Figure~\ref{fig:contact_gen}--a); not too close, because the robot must avoid collision.
We approximate $C_{Contact}^0$ with a set $C_{reach}^0$, defined by the \textit{reachability condition}, a geometrical criterion based on this intuitive description (Figure~\ref{fig:dedefeas} -- Red).
Our objective is to find a root path such that the root (scaled by a factor $s \geq 1$) is not in collision, while the reachable workspace is in collision with the environment.
In Section~\ref{rbprm} we detail how  $C_{reach}^0$ is designed and sampled to compute a guide path with a dedicated planner, the Reachability-Based RRT---RB-RRT--- (Figure~\ref{fig:framework}---\Pa).

@@

The second stage extends the guide path into a sequence of contact configurations (Figure~\ref{fig:framework}--\Pb).
With a fixed root position, the dimensionality of the 
contact generation problem is reduced to the number of degrees of freedom of the considered limb. In other words, we consider each limb as a manipulator attached to the root, and select the most relevant contact from a database of precomputed configurations, which is independent from the environment (Figure~\ref{fig:contact_gen}).

@@

There is a combinatorial issue in the choice of which contact to create,
and where to create it. This complexity is avoided by selecting the ``most relevant'' contact configurations with user-defined heuristics. In Section~\ref{sec:heuristics} we present extensively those heuristics that have been used in our experiments.

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + r$} degrees of freedom (DOFs), with $r \geq 6$. $r$ denotes the world coordinates of the root of the robot,
as well as extra DOFs not related to the limbs (For instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$. 
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to a root.
It is described by a configuration $\mathbf{q} \in  \mathbb{R}^r \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ denotes the position and orientation of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the trunk of the robot is denoted $W^0$
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$ %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.

>>>>

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed in a sequential fashion: we first plan a root guide path with the \textit{reachability condition}, before
extending it into a sequence of configurations in static equilibrium.
In this Section we describe this workflow and give some notations used throughout the paper. Some key concepts are introduced,
whose definitions can be found in the glossary at the end of this paper.

@@

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a root guide path. 
%The objective is to compute a path of root configurations that will allow for equilibrium. 
%A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of configurations in static equilibrium.
\begin{figure}[t]
\centering
  \begin{overpic}[width=1\linewidth]{figures/2D_feas}
		\put (15,) {$C_{Equil}^0$      $\subset$} 
		\put (47,) {$C_{Contact}^0$ $\approx$ } 
		\put (76,) {$C_{Reach}^0$} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Illustration of several root configurations sets used in this paper in a 2D scene. Obstacles are violet, and units are in meters. To show the sets in a 2D representation, all the rotational joints of HRP-2 are locked in the shown configuration, such that a torso configuration
is only described by two positional parameters (x and y). $C_{Equil}^0$ is included in $C_{Contact}^0$. $C_{Reach}^0$ approximates $C_{Contact}^0$. Depending on a parametrization, we can obtain $C_{Contact}^0 \subset C_{Reach}^0$. Considering the configurations around the top obstacle, we can observe a dramatic
divergence between  $C_{Equil}^0$  and $C_{Contact}^0$ when the problem is not \textit{\gls{cluttered}}.}
		   \label{fig:dedefeas}
\end{figure}
%
Ideally we would like to sample root configurations that are \gls{equilibrium feasible}. A root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that:
\begin{description}
\item[C1]: the robot is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: the robot is in static equilibrium. 
\end{description}
Verifying C2 is computationally expensive because it requires the projection of the whole-body configuration into a contact posture, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.
For this reason, we are interested in finding an approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} that is computationally efficient.
A requirement for C2 is the condition:
\begin{description}
\item[C3]: at least one end-effector is in contact with the environment.
\end{description}
C1 and C3 define \glslink{contact feasible}{\textit{contact feasibility}}, which is easier to verify than \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} because it does not require the computation of contact forces.
We make the assumption that for the \textit{\gls{cluttered} contact planning problem}, most of the times \gls{contact feasible} root configurations are also \gls{equilibrium feasible}.
For this reason we neglect C2 during the resolution of \Pa, and we consider it only when solving \Pb.
However, even verifying \glslink{contact feasible}{\textit{contact feasibility}} is too computationally expensive because it requires the projection of the whole-body configuration into a contact posture.

An intuitive description of \gls{contact feasible} configurations is ``close, but not too close'': close, because a contact surface must be partially included in the reachable workspace of the robot (represented for the right leg in Figure~\ref{fig:contact_gen}--a); not too close, because the robot must avoid collision.
We approximate \glslink{contact feasible}{\textit{contact feasibility}} with the \textit{reachability condition}, a geometrical criterion based on this intuitive description.
More precisely, a root configuration is \textit{reachable} if the root scaled by a factor $s \geq 1$ is not in collision, while the reachable workspace is in collision with the environment.
Verifying the \textit{reachability condition} is very efficient in practice.
In Section~\ref{rbprm} we detail how  we define and use the \textit{reachability condition} to compute a guide path with the Reachability-Based RRT (Figure~\ref{fig:framework}---\Pa).

Figure~\ref{fig:dedefeas} gives an insight into the difference between the three conditions, depicting $C_{Equil}^0$, $C_{Contact}^0$ and $C_{Reach}^0$, which are the sets of \gls{equilibrium feasible}, \gls{contact feasible} and \textit{reachable} root configurations, respectively.

@@

The second stage extends the guide path into a sequence of contact configurations (Figure~\ref{fig:framework}--\Pb).
With a fixed root position, the dimensionality of the 
contact generation problem is reduced to the number of degrees of freedom of the considered limb. 
%In other words, we consider each limb as a manipulator attached to the root. 
We select from a database of precomputed limb configurations (independent from the environment) the ones resulting in the end-effector being close to the environment.
Limiting our selection to this set of precomputed configurations allow us to drastically reduce the computation times. 
Then we select one of these configurations based on user-defined heuristics (Figure~\ref{fig:contact_gen}), which we present in Section~\ref{sec:heuristics}.

@@

@@

\medskip
\textbf{A robot} is a kinematic chain $R$, comprising \mbox{$n + r$} degrees of freedom (DOFs), with $r \geq 6$. $r$ denotes the world coordinates of the root of the robot,
as well as extra DOFs not related to the limbs (for instance, HRP-2 has two extra DOFS in the torso, such that we have $r=8$). 
$R$ is composed of $l$ limbs $R^k, 1 \leq k \leq l$, attached to a root.
It is described by a configuration $\mathbf{q} \in  \mathbb{R}^r \times \mathbb{R}^n$.
We define some relevant projections of $\mathbf{q}$:
\begin{itemize}
	\item $\mathbf{q}^k$ denotes the configuration (a vector of joint values) of the limb $R^k$; % (Figure~\ref{fig:character});
	\item $\mathbf{q}^{\overline{k}}$ denotes the vector of joint values of R \textbf{not} related to $R^k$. We define for convenience \mbox{$\mathbf{q}= \mathbf{q}^k \oplus \mathbf{q}^{\overline{k}}$}; %Abusively, we denote by $\mathbf{q}^0 \oplus \mathbf{q}^k$ any configuration respecting $\mathbf{q}^0$ and $\mathbf{q}^k$;
	\item $\mathbf{q}^{0}\in \mathbb{R}^r$ denotes the world coordinates of the root of the robot $R$.
\end{itemize}
%Finally, $ \mathbf{q}_{start}$ and  $\mathbf{q}_{goal}$ are the start and goal configurations given as an input of our problem.

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing the trunk of the robot is denoted $W^0$.
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\},
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define $W = \bigcup_{k=1}^{l}W^k$, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ translated and rotated by the rigid displacement~$\mathbf{q}^{0}$.
